---
title: "Raised Incidence Model for Hamburg Case"
geometry: "left=2cm,right=2cm,top=1cm,bottom=1cm"
output:
  pdf_document: 
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
Sys.setenv(lang = "en_US")
```

This script implements the raised incidence model by Diggle (1990) and the extensions to accommodate multiple potential sources by Diggle & Rowlingson (1994) on Data of the region of Hamburg.

-   Population Data as the population at risk [$\lambda_0(x)$]

-   Retailer location data as sources [$x_k$]

-   Outbreak case locations as outbreak [$\lambda_1(x)$]

All spatial data was initially stored in the coordinate system ETRS89-LAEA Europe - EPSG:3035 and transformed from metres to kilometres.

# Load Libraries

```{r message=FALSE}
library(spatstat)
library(sf)
```

# Load Outbreak from Shapefile

```{r}
sf_outbreak <- st_read("./Data/Outbreaks/Aldi_20/Aldi_Outbreak.shp")
```


```{r eval=FALSE, include=FALSE}
plot(st_geometry(sf_outbreak), pch = 19, main = "Artificial Outbreak")
```

**Convert into spatstat ppp object**

```{r warning=FALSE}
ppp_outbreak <- as.ppp(X=sf_outbreak$geometry, W=owin(c(4303150,4342650), c(3365250,3403550)))
ppp_outbreak
```
```{r}
ppp_outbreak <- rescale(ppp_outbreak, 1000, "km")
```

  
**Plot**

```{r}
plot(ppp_outbreak, pch = 19,  main = "Artificial Outbreak")
```

# Load Population Data

We applied kernel smoothing (Diggle, 1985) to the population data to obtain an unnormalised estimate of the spatially varying population density of susceptibles ($\lambda_0(x)$).

```{r}
im_population <- readRDS("./Data/Population Data/im_population.rds")
```

**Plot**

```{r}
colors <- c("#ffffff", "#FEEDDE", "#FDBE85", "#FD8D3C", "#E6550D","#F01B00", "#A63603", "#732600")
plot(im_population, main = "Kernel Smoothed Population Density Hamburg 2011", col =colors)
contour(im_population, main = "Contour of Population Density Hamburg 2011")
```

# Load Potential Pattern

For now, we only use the one point source model, so we read only the Shapefile of one store of the chain.

```{r}
sf_stores <- st_read("./Data/Potential Pattern/Aldi_store_6.shp")
```

```{r eval=FALSE, include=FALSE}
plot(st_geometry(sf_stores), pch= 7, main = "Source")
```

**Convert into spatstat ppp object**

```{r}
ppp_stores <- as.ppp(X=sf_stores$geometry, W=owin(c(4303150,4342650), c(3365250,3403550)))
ppp_stores
```
```{r}
ppp_stores <- rescale(ppp_stores, 1000, "km")
```

**Plot**

```{r}
plot(ppp_stores, pch = 7, main = "Source")
```

```{r}
print(ppp_stores$x[1])
print(ppp_stores$y[1])
```

# Fit Null Model

$H_0: \lambda_1(x) = \rho \lambda_0(x)$

```{r}
fit0 <- ppm(ppp_outbreak ~ offset(log(im_population)), gcontrol = glm.control(maxit = 10000))
```

```{r}
fit0
```

# Fit Alternative Model

For one source: $H_0: \lambda_1(x) = \rho\lambda_0(x) (1+ \alpha \ e^{-\beta x'x})$

We define the squared distance to the source:

```{r}
d2source <- function(x, y, xPOS= 4316.914, yPOS=3385.953) {
  (x - xPOS)^2 + (y - yPOS)^2}
```

We define the functional form of the raised incidence part:

```{r}
raisin <- function(x, y, alpha, beta) {
  1 + alpha * exp(-beta * d2source(x, y))
}
```

We fit the model to our outbreak



```{r}
fit1 <- ippm(ppp_outbreak ~ offset(log(im_population) + log(raisin)),
  start = list(alpha = 5, beta = 1), gcontrol = glm.control(maxit = 1000))
```


```{r}
fit1
```




## Statistical Inference comparing the two models

```{r}
anova(fit0, fit1, test="LRT")
```

# Quadscheme
```{r}
Q <- quadscheme(ppp_outbreak, eps = 100)
```

```{r}
fit0_test <- ppm(Q ~ offset(log(im_population)), gcontrol = glm.control(maxit = 10000))
fit0_test
```

```{r}
fit1_test <- ippm(Q ~ offset(log(im_population) + log(raisin)),
  start = list(alpha = 5, beta = 1), gcontrol = glm.control(maxit = 1000))
fit1_test
```

# Multiple Sources
```{r}
all_raisins = list()
for (i in 1:3){#length(ppp_stores)  {
    all_raisins[i] =  paste0("(1 + alpha * exp(-beta * ((x- ",ppp_stores$x[i],")^2 + (y- ",ppp_stores$y[i],")^2)))")
}

raisin_func <- function(x, y, alpha, beta) {
  paste(all_raisins, collapse = "*")
}



```




```{r}
d2source <- function(x, y, xPOS, yPOS) {
  (x - xPOS)^2 + (y - yPOS)^2}
```

We define the functional form of the raised incidence part:

```{r}
raisin <- function(x, y, alpha, beta) {
  1 + alpha * exp(-beta * d2source(x, y))
}
```